### 2021-03-15 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Optimized chapters 3, 4 and 7

\begin{table}[!t]
  \centering
  \csvreader[ECSAndTCRTable]{
    ch04_papers_ecs_tcr_assessment/data/cmip5_ecs-tcr.csv}{}{
    \dataset & $\idx$ & $\ecs$ & $\tcr$}
  \caption[
    \Acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models.
  ]{
    \Acf{ECS} and \acf{TCR} evaluated for the \acs{CMIP}5 models. Details on
    the calculation of \acs{ECS} and \acs{TCR} are given in
    \cref{subsec:02:ecs} and \cref{subsec:02:tcr}, respectively. The \acf{MMM}
    is calculated from the Gregory regression method using the \acs{MMM} net
    \acf{TOA} radiation and the \acs{MMM} change in \acf{GSAT} similar to
    \cref{fig:04:gregory_regression_cmip6_mmm}. The multi-model standard
    deviation is given by the sample standard deviation of \acs{ECS} evaluated
    over all climate models (using the normalization $1 / M$, where $M$ is the
    number of models). Corresponding references for each model are given in
    \cref{tab:app:a:cmip5_models}. \AdaptedFrom{Meehl2020}.
  }
  \label{tab:04:ecs_tcr_cmip5}
\end{table}

@@

To illustrate the results of the previous
\namecref{sec:04:evaluation_ecs_and_tcr} and put them into historical context,
\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
and the different \acp{AR} of the \ac{IPCC} in combination with the
corresponding modeled ranges from the different \ac{CMIP} generations. Since
the Charney report in 1979, the assessed range of \ac{ECS} of
$\rangeunit{1.5}{4.5}{K}$ has almost remained unchanged during the last 40
years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
  Stocker2013} with the exception of \acs{AR}4, where the lower bound was
temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
\acs{AR}4 and \acs{AR}5, the corresponding climate model generations from
\acs{CMIP}1, \acs{CMIP}3 and \acs{CMIP}5 more or less agree with this assessed
range \autocite{Kattenberg1996, Randall2007, Flato2013}. In contrast to that,
some of the early climate models used in \acs{AR}1 and some \acs{CMIP}2 models
used in \acs{AR}3 exhibit values of \acs{ECS} well above $4.5 \unit{K}$,
resulting in a upper model range of about $0.6 \unit{K}$ higher than the
assessed range \autocite{Mitchell1990, Cubasch2001}. However, these deviations
from the assessed range are insignificant to the ones present in the
\acs{CMIP}6 ensemble, which have become evident over the course of 2020. As
shown in \cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper
bound are more extreme than in any previous climate model generation, resulting
in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$ (as of
\TheMonth{}). This increase in range is particularly relevant for the upper
bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that
exceeds this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third
(14 models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to our analysis,
several modeling centers independently confirmed high \ac{ECS} values in their
models \autocite{Andrews2019, Gettelman2019, Wyser2020}. A list of possible
reasons for the increased climate sensitivity in the \acs{CMIP}6 ensemble
including aerosol-cloud interactions and changes in the shortwave cloud
feedback over the Southern Ocean is given in
\cref{sec:04:possible_reasons_high_ecs_cmip6}.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch04_papers_ecs_tcr_assessment/figs/cloud_phase_change_feedback.pdf}
  \caption[
    Schematic illustration of the strong negative shortwave cloud phase change
    feedback over the Southern Ocean.
  ]{
    Schematic illustration of the strong negative shortwave cloud phase change
    feedback over the Southern Ocean, which is present in earlier \acs{CMIP}
    model generations like \acs{CMIP}5 (green arrows). Due to substantial
    improvements in the microphysical representation of mixed-phase clouds in
    many \acs{CMIP}6 models, the predominantly ice clouds over the Southern
    Ocean in the present-day climate of previous climate model generations have
    been replaced with clouds that predominantly consist of (supercooled)
    liquid clouds \autocite{BodasSalcedo2019, Gettelman2019, Mulcahy2020,
      Williams2020}. Thus, in these models a cloud phase change due to global
    warming is no longer possible, which removes (or significantly reduces) the
    strong negative cloud phase change feedback (orange arrows). This leads to
    an increase in the global net cloud feedback \autocite{BodasSalcedo2019,
      Tan2016} and the \acf{ECS} in the corresponding \acs{CMIP}6 models
    \autocite{Andrews2019, Gettelman2019}, which is a possible explanation for
    the high \acs{ECS} values found in many \acs{CMIP}6 models.
  }
  \label{fig:04:cloud_phase_change_feedback}
\end{figure}

However, it is well known that the low-level mixed-phase clouds over the
Southern Ocean in these earlier climate models are biased towards too large
amounts of ice crystals and too little amounts of supercooled liquid water
\autocite{BodasSalcedo2016, McCoy2015, McCoy2016, Tan2016}. For this reason,
the microphysical representation of mixed-phase clouds has been substantially
improved in many \acs{CMIP}6 models \autocite{BodasSalcedo2019, Gettelman2019,
  Mulcahy2020, Williams2020}, leading for example to a better agreement of the
simulated \ac{TOA} shortwave \ac{CRE} with corresponding observations
\autocite{Bock2020}. As a consequence, the strong negative cloud feedback
described above that results from a cloud phase change from ice clouds in the
present-day to liquid clouds in the future is removed (or significantly
reduced) in those \acs{CMIP}6 models that simulate predominantly liquid clouds
over the present-day Southern Ocean. This is illustrated in
\cref{fig:04:cloud_phase_change_feedback} (orange arrows). Moreover, this
reduction of the negative shortwave cloud phase change feedback increases the
global net cloud feedback \autocite{BodasSalcedo2019, Tan2016} and eventually
the \ac{ECS} of the affected \acs{CMIP}6 models \autocite{Andrews2019,
  Gettelman2019}, offering a possible explanation for the high \ac{ECS} values
in some \acs{CMIP}6 models.

@@

To further confirm this hypothesis, we show the global distributions of
relevant cloud feedback parameters (expressed by the net, shortwave and
longwave \ac{CRE} feedback parameters; see
\cref{subsec:02:cloud_feedback_parameters}) for the \acs{CMIP}6 \ac{MMM} and
its difference to the \acs{CMIP}5 \ac{MMM} in
\cref{fig:04:cloud_feedback_parameters}. First of all, the comparison of
\cref{fig:04:cloud_feedback_parameters:a,fig:04:cloud_feedback_parameters:c,%
  fig:04:cloud_feedback_parameters:e} demonstrates that the shortwave component
dominates the net \ac{CRE} feedback parameter over large swaths of the globe
in the \acs{CMIP}6 \ac{MMM}. The sign change at approximately $60
\unit{\degree S}$ in the shortwave \ac{CRE} feedback parameter (see
\cref{fig:04:cloud_feedback_parameters:b}) shows where the climate models are
switching from simulating almost purely liquid clouds (north of $60
\unit{\degree S}$) to clouds with an increasing ice component (south of $60
\unit{\degree S}$) in their pre-industrial and present-day experiments. The
increasing fraction of ice in the simulated mixed-phase clouds when moving
south supports a negative shortwave cloud feedback on warming in these areas.
\Cref{fig:04:cloud_feedback_parameters:b,fig:04:cloud_feedback_parameters:d}
show a higher shortwave and net \ac{CRE} feedback parameter over the Southern
Ocean in \acs{CMIP}6 compared to \acs{CMIP}5. This is consistent to the
hypothesis presented in the paragraph above, which claims that the reduced
negative cloud phase change feedback over the Southern Ocean in \acs{CMIP}6
leads to an increased net cloud feedback parameter in comparison to older
\ac{CMIP} generations. This result is further supported by
\textcite{Zelinka2020}, who find a larger positive zonal mean net cloud
feedback parameter in \acs{CMIP}6 compared to \acs{CMIP}5 due to a more
positive (reduced negative) low-level cloud feedback, primarily in the
extratropics. In comparison to earlier \acs{CMIP} generations, the
\acs{CMIP}6 models exhibit weaker increases in the extratropical low-level
cloud cover and the corresponding liquid water content as a result of global
warming. Moreover, \textcite{Zelinka2020} show that this mainly results from
an increase in the liquid condensate fraction in the \acs{CMIP}6 clouds for
the pre-industrial and present-day periods, leading to the aforementioned
reduction in the negative cloud phase change feedback on warming and
eventually to higher \ac{ECS} values.

>>>>

\begin{table}[!t]
  \centering
  \csvreader[ECSAndTCRTable]{
    ch04_papers_ecs_tcr_assessment/data/cmip5_ecs-tcr.csv}{}{
    \dataset & $\idx$ & $\ecs$ & $\tcr$}
  \caption[
    \Acf{ECS} and \acf{TCR} calculated for the \acs{CMIP}5 models.
  ]{
    \Acf{ECS} and \acf{TCR} calculated for the \acs{CMIP}5 models. Details on
    this are given in \cref{subsec:02:ecs} and \cref{subsec:02:tcr},
    respectively. The \acf{MMM} is calculated from the Gregory regression
    method using the \acs{MMM} net \acf{TOA} radiation and the \acs{MMM} change
    in \acf{GSAT} similar to \cref{fig:04:gregory_regression_cmip6_mmm}. The
    multi-model standard deviation is given by the sample standard deviation of
    \acs{ECS} evaluated over all climate models (using the normalization $1 /
    M$, where $M$ is the number of models). Corresponding references for each
    model are given in \cref{tab:app:a:cmip5_models}. \AdaptedFrom{Meehl2020}.
  }
  \label{tab:04:ecs_tcr_cmip5}
\end{table}

@@

To illustrate the results of the previous
\namecref{sec:04:evaluation_ecs_and_tcr} and put them into historical context,
\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
and the different \acp{AR} of the \ac{IPCC} in combination with the
corresponding modeled ranges from the different \ac{CMIP} generations. Since
the Charney report in 1979, the assessed range of \ac{ECS} of
$\rangeunit{1.5}{4.5}{K}$ has almost remained unchanged during the last 40
years \autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
  Stocker2013} with the exception of \acs{AR}4, where the lower bound was
temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
\acs{AR}4 and \acs{AR}5, the corresponding climate model generations from
\acs{CMIP}1, \acs{CMIP}3 and \acs{CMIP}5 more or less agree with this assessed
range \autocite{Kattenberg1996, Randall2007, Flato2013}. In contrast to that,
some of the early climate models used in \acs{AR}1 and some \acs{CMIP}2 models
used in \acs{AR}3 exhibit values of \acs{ECS} well above $4.5 \unit{K}$,
resulting in a upper model range of about $0.6 \unit{K}$ higher than the
assessed range \autocite{Mitchell1990, Cubasch2001}. However, these deviations
from the assessed range are small compared to the ones present in the
\acs{CMIP}6 ensemble, which have become evident over the course of 2020. As
shown in \cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper
bound are more extreme than in any previous climate model generation, resulting
in a \acs{CMIP}6 \ac{ECS} model range of $\rangeunit{1.8}{5.6}{K}$ (as of
\TheMonth{}). This increase in range is particularly relevant for the upper
bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that
exceeds this upper bound: out of the 42 analyzed \acs{CMIP}6 models, a third
(14 models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to our analysis,
several modeling centers independently confirmed high \ac{ECS} values in their
models \autocite{Andrews2019, Gettelman2019, Wyser2020}. A list of possible
reasons for the increased climate sensitivity in the \acs{CMIP}6 ensemble
including aerosol-cloud interactions and changes in the shortwave cloud
feedback over the Southern Ocean is given in
\cref{sec:04:possible_reasons_high_ecs_cmip6}.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch04_papers_ecs_tcr_assessment/figs/cloud_phase_change_feedback.pdf}
  \caption[
    Schematic illustration of the strong negative shortwave cloud phase change
    feedback over the Southern Ocean.
  ]{
    Schematic illustration of the strong negative shortwave cloud phase change
    feedback over the Southern Ocean, which is present in earlier \acs{CMIP}
    model generations like \acs{CMIP}5 (green arrows). Due to substantial
    improvements in the microphysical representation of mixed-phase clouds in
    many \acs{CMIP}6 models, the predominantly ice clouds over the Southern
    Ocean in the present-day climate of previous climate model generations have
    been replaced with clouds that predominantly consist of (supercooled)
    liquid clouds \autocite{BodasSalcedo2019, Gettelman2019, Mulcahy2020,
      Williams2020}. Thus, in the affected \acs{CMIP}6 models, this cloud phase
    change due to warming is reduced by allowing for supercooled cloud liquid,
    which substantially reduces the strong negative cloud phase change feedback
    (orange arrows). This leads to an increase in the global net cloud feedback
    \autocite{BodasSalcedo2019, Tan2016} and the \acf{ECS} in the corresponding
    \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019}, which is a
    possible explanation for the high \acs{ECS} values found in many
    \acs{CMIP}6 models.
  }
  \label{fig:04:cloud_phase_change_feedback}
\end{figure}

However, it is well known that the low-level mixed-phase clouds over the
Southern Ocean in these earlier climate models are biased towards too large
amounts of ice crystals and too little amounts of supercooled liquid water
\autocite{BodasSalcedo2016, McCoy2015, McCoy2016, Tan2016}. For this reason,
the microphysical representation of mixed-phase clouds has been substantially
improved in many \acs{CMIP}6 models \autocite{BodasSalcedo2019, Gettelman2019,
  Mulcahy2020, Williams2020}, leading for example to a better agreement of the
simulated \ac{TOA} shortwave \ac{CRE} with corresponding observations
\autocite{Bock2020}. As a consequence, the strong negative cloud feedback
described above that results from a cloud phase change from ice clouds in the
present-day to liquid clouds in the future is substantially reduced in those
\acs{CMIP}6 models that simulate predominantly liquid clouds over the
present-day Southern Ocean. This is illustrated in
\cref{fig:04:cloud_phase_change_feedback} (orange arrows). Moreover, this
reduction of the negative shortwave cloud phase change feedback increases the
global net cloud feedback \autocite{BodasSalcedo2019, Tan2016} and eventually
the \ac{ECS} of the affected \acs{CMIP}6 models \autocite{Andrews2019,
  Gettelman2019}, offering a possible explanation for the high \ac{ECS} values in
some \acs{CMIP}6 models.

@@

To further confirm this hypothesis, we show the global distributions of
relevant cloud feedback parameters (expressed by the net, shortwave and
longwave \ac{CRE} feedback parameters; see
\cref{subsec:02:cloud_feedback_parameters}) for the \acs{CMIP}6 \ac{MMM} and
its difference to the \acs{CMIP}5 \ac{MMM} in
\cref{fig:04:cloud_feedback_parameters}. First of all, the comparison of
\cref{fig:04:cloud_feedback_parameters:a,fig:04:cloud_feedback_parameters:c,%
  fig:04:cloud_feedback_parameters:e} demonstrates that the shortwave component
dominates the net \ac{CRE} feedback parameter over large swaths of the globe
in the \acs{CMIP}6 \ac{MMM}. The sign change at approximately $60
\unit{\degree S}$ in the shortwave \ac{CRE} feedback parameter (see
\cref{fig:04:cloud_feedback_parameters:c}) roughly shows where the climate
models simulate a high fraction of ice clouds (south of $60 \unit{\degree
  S}$) for which possible phase changes under global warming are potentially
important. Such possible phase changes (from ice to liquid) contribute to a
negative shortwave cloud feedback in this region as liquid clouds tend to be
more reflective than ice clouds with a similar cloud water content.
\Cref{fig:04:cloud_feedback_parameters:b,fig:04:cloud_feedback_parameters:d}
show a higher shortwave and net \ac{CRE} feedback parameter over the Southern
Ocean in \acs{CMIP}6 compared to \acs{CMIP}5. This is consistent to the
hypothesis presented in the paragraph above, which claims that the reduced
negative cloud phase change feedback over the Southern Ocean in \acs{CMIP}6
leads to an increased net cloud feedback parameter in comparison to older
\ac{CMIP} generations. This result is further supported by
\textcite{Zelinka2020}, who find a larger positive zonal mean net cloud
feedback parameter in \acs{CMIP}6 compared to \acs{CMIP}5 due to a more
positive (reduced negative) low-level cloud feedback, primarily in the
extratropics. In comparison to earlier \acs{CMIP} generations, the
\acs{CMIP}6 models exhibit weaker increases in the extratropical low-level
cloud cover and the corresponding liquid water content as a result of global
warming. Moreover, \textcite{Zelinka2020} show that this mainly results from
an increase in the liquid condensate fraction in the \acs{CMIP}6 clouds for
the pre-industrial and present-day periods, leading to the aforementioned
reduction in the negative cloud phase change feedback on warming and
eventually to higher \ac{ECS} values.

### 2021-03-15 ch07_conclusion/conclusion.tex
Optimized chapters 3, 4 and 7

The analysis of future climate projections from numerical climate model
simulations is of paramount importance for an accurate quantification of the
future climate change. Since this involves metrics of public interest like the
allowable fossil fuel emissions to meet particular warming targets, for example
the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015},
this scientific question is not only relevant to climate scientists but to
policymakers and the whole human society. In the light of the large spread of
climate sensitivity in the most recent generation of climate models from
\acs{CMIP}6 \autocite{Meehl2020}, a careful statistical evaluation and
refinement of the output of multi-model climate projections is as relevant as
ever. This thesis quantifies the associated uncertainties, presents the
evaluation of established methods to reduce these uncertainties in the new
climate model ensemble and finally introduces a novel alternative technique
based on supervised \ac{ML}.

@@

In a first analysis, the climate sensitivity metrics \ac{ECS} and \ac{TCR} are
evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved simulation of the shortwave \ac{CRE} in these
models when compared to observations. However, this change also substantially
reduces the strong negative shortwave cloud feedback over the Southern Ocean
that is present in previous \ac{CMIP} generations and that results from a cloud
phase change from ice clouds in the present-day to liquid clouds in the future.
In the affected \acs{CMIP}6 models, this cloud phase change is not possible
anymore since the models predominantly simulate liquid clouds over the
present-day Southern Ocean.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, suitable emergent constraints on the \acs{CMIP}6 ensemble are
evaluated. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. In total eleven emergent constraints on \ac{ECS} are
assessed, which are mostly related to cloud feedbacks since these constitute
the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
  Flato2013}. Since all of the evaluated emergent constraints have been derived
on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, out-of-sample tests on the emergent
constraints can be performed by assessing whether they still hold for the
\acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
already published in \textcite{Schlund2020a}, a substantial reduction of skill
for the majority of emergent constraints is found when applied to the
\acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
skill is expressed as a decrease of the coefficient of determination $R^2$ of
the emergent relationship and a decrease of the statistical significance using
the null hypothesis that there is no correlation between the predictor and
\ac{ECS}. Moreover, the corresponding \acp{PDF} for the emergent constraints
show higher $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
confidence) for almost all emergent constraints, resulting in values of
$\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6 (\acs{CMIP}5:
$\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent constraints, this is
an increase in the $66 \unit{\%}$ \ac{ECS} range of $51 \unit{\%}$. Similarly,
the best estimates for \ac{ECS} show values of $\rangeunit{2.97}{3.88}{K}$ in
\acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in \acs{CMIP}6, resulting in an
increase of about $12 \unit{\%}$ averaged over all emergent constraints. Thus,
\cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
be answered with a \enquote{no} for the \acs{CMIP}6 ensemble. The increased
best estimates and spreads resulting from the emergent constraints in
\acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
of \ac{ECS} in \acs{CMIP}6. A possible reason for the reduced skill of the
emergent constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

@@

\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for the
increased climate sensitivity in the \acs{CMIP}6 models (changes in the
aerosol-cloud interaction and changes in the shortwave cloud phase change
feedback over the Southern Ocean). While these are likely explanations, there
might be additional relevant aspects that have not been analyzed in detail yet.
This could include additional feedback processes that are not present in older
\ac{CMIP} generations or missing feedback processes that appear in previous
\ac{CMIP} models but are absent in the \acs{CMIP}6 ensemble. Obvious candidates
for these feedback processes are cloud-related feedbacks, which pose a major
source of uncertainty in climate sensitivity in modern-day \acp{ESM}
\autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms
can further help to gain a better understanding of the \acs{CMIP}6 models and
potentially to reduce associated uncertainties in the entire multi-model
ensemble.

Due to the flexibility of the new \ac{ML}-based climate model weighting
approach presented in \cref{ch:06:paper_gpp}, one natural way forward is to
apply this technique to other target variables. While the scalar metrics
\ac{ECS} or \ac{TCR} with one value per climate model do not offer enough
training points for this technique, a possible alternative could be the gridded
\acl{T} over the \nth{21} century in the \ac{SSP} scenarios. A relevant study
in this context is given by \textcite{Brunner2020}, who use performance and
interdependence-based climate model weighting to constrain the \acl{T} over the
\nth{21} century based on observable process-based predictors of today's
climate and which could serve as a possible baseline for the new \ac{ML}-based
weighting approach. An exciting and valuable addition to the method could be
causal inference \autocite{Nowack2020, Runge2019}. In the current approach, the
relationships between the different predictors and the target variable are
based on statistical correlation alone. Moreover, the local feature importance
maps do not reveal true causal relationships between the predictors and the
target variable, but only show the relative weight that is given to a specific
predictor at a specific location for the prediction of the target variable by
the \ac{ML} model. By integrating causal networks into the new approach, it
might be possible to explore and utilize true causal connections instead.

>>>>

The analysis of future climate projections from numerical climate model
simulations is of paramount importance for an accurate quantification of the
future climate change. Since this involves metrics of public interest like the
allowable fossil fuel emissions to meet particular warming targets, for example
the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement \autocite{UNFCCC2015},
research in this field of science is not only relevant to climate scientists
but to policymakers and the whole human society. In the light of the large
spread of climate sensitivity in the most recent generation of climate models
from \acs{CMIP}6 \autocite{Meehl2020}, a careful statistical evaluation and
refinement of the output of multi-model climate projections is as relevant as
ever. This thesis quantifies the associated uncertainties, presents the
evaluation of established methods to reduce these uncertainties in the new
climate model ensemble and finally introduces a novel alternative technique
based on supervised \ac{ML}.

@@

In a first analysis, the climate sensitivity metrics \ac{ECS} and \ac{TCR} are
evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved simulation of the shortwave \ac{CRE} in these
models when compared to observations. However, this change also substantially
reduces the strong negative shortwave cloud feedback over the Southern Ocean
that is present in previous \ac{CMIP} generations and that results from a cloud
phase change from ice clouds in the present-day to liquid clouds in the future.
In the affected \acs{CMIP}6 models, this cloud phase change due to warming is
reduced by allowing for supercooled cloud liquid. As a consequence, these
models simulate less cloud ice over the present-day Southern Ocean than their
predecessor versions with no supercooled cloud liquid.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, suitable emergent constraints on the \acs{CMIP}6 ensemble are
evaluated. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. In total eleven emergent constraints on \ac{ECS} are
assessed, which are mostly related to cloud feedbacks since these constitute
the most important source of uncertainty for \ac{ECS} \autocite{Boucher2013,
  Flato2013}. Since all of the evaluated emergent constraints have been derived
on the \acs{CMIP}3 or \acs{CMIP}5 ensemble, out-of-sample tests on the emergent
constraints can be performed by assessing whether they still hold for the
\acs{CMIP}6 models. In this study, which is shown in \cref{ch:05:paper_ecs} and
already published in \textcite{Schlund2020a}, a substantial reduction of skill
for the majority of emergent constraints is found when applied to the
\acs{CMIP}6 ensemble in comparison to the \acs{CMIP}5 ensemble. This drop in
skill is expressed as a decrease of the coefficient of determination $R^2$ of
the emergent relationship and a decrease of the statistical significance using
the null hypothesis that there is no correlation between the predictor and
\ac{ECS}. Moreover, the corresponding \acp{PDF} for the emergent constraints
show higher $66 \unit{\%}$ \ac{ECS} ranges ($\rangeunit{17}{83}{\%}$
confidence) for almost all emergent constraints, resulting in values of
$\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6 (\acs{CMIP}5:
$\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent constraints, this is
an increase in the $66 \unit{\%}$ \ac{ECS} range of $51 \unit{\%}$. Similarly,
the best estimates for \ac{ECS} show values of $\rangeunit{2.97}{3.88}{K}$ in
\acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in \acs{CMIP}6, resulting in an
increase of about $12 \unit{\%}$ averaged over all emergent constraints. Thus,
\cref{enum:01:question_2} (\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to
be answered with a \enquote{not very well} for the \acs{CMIP}6 ensemble. The
increased best estimates and spreads resulting from the emergent constraints in
\acs{CMIP}6 are likely related to the increased \ac{MMM} and multi-model spread
of \ac{ECS} in \acs{CMIP}6. A possible reason for the reduced skill of the
emergent constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

@@

Climate sensitivity is a policy-relevant and easy-to-use metric to assess the
strength of climate change. The range of model results for this important
climate metric, however, remains large and has not been narrowed down over the
last decades. In the latest generation of climate models contributing to
\acs{CMIP}6, the range has even increased.
\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for this:
changes in the aerosol-cloud interaction and changes in the shortwave cloud
phase change feedback over the Southern Ocean. While these are likely
explanations, there might be additional relevant aspects that have not been
analyzed in detail yet. This could include additional feedback processes that
are not present in older \ac{CMIP} generations or missing feedback processes
that appear in previous \ac{CMIP} models but are absent in the \acs{CMIP}6
ensemble. Obvious candidates for these feedback processes are cloud-related
feedbacks, which pose a major source of uncertainty in climate sensitivity in
modern-day \acp{ESM} \autocite{Boucher2013}. Identifying and quantifying such
feedback mechanisms can further help to gain a better understanding of the
\acs{CMIP}6 models and potentially to reduce associated uncertainties in the
entire multi-model ensemble.

A promising way forward to reduce uncertainties in multi-model climate
projections is to apply the new flexible \ac{ML}-based climate model weighting
approach introduced in \cref{ch:06:paper_gpp} to other target variables. While
the scalar metrics \ac{ECS} or \ac{TCR} with one value per climate model do not
offer enough training points for this technique, a possible alternative could
be the gridded \acl{T} over the \nth{21} century in the \ac{SSP} scenarios. A
relevant study in this context is given by \textcite{Brunner2020}, who use
performance and interdependence-based climate model weighting to constrain the
\acl{T} over the \nth{21} century based on observable process-based predictors
of today's climate and which could serve as a possible baseline for the new
\ac{ML}-based weighting approach. An exciting and valuable addition to the
method could be causal inference \autocite{Nowack2020, Runge2019}. In the
current approach, the relationships between the different predictors and the
target variable are based on statistical correlation alone. Moreover, the local
feature importance maps do not reveal true causal relationships between the
predictors and the target variable, but only show the relative weight that is
given to a specific predictor at a specific location for the prediction of the
target variable by the \ac{ML} model. By integrating causal networks into the
new approach, it might be possible to explore and utilize true causal
connections instead.
